---
title: "Racial Disparities and Trends in Endometrial Cancer"
author: "A McVey Moffatt"
date: "`r Sys.Date()`"
output: html_document
---

### Introduction

Endometrial cancer (EC) is the abnormal proliferation of cells in the inner epithelial lining of the uterus. EC is the fourth most common cancer in women and most common female genital tract cancer. The objective of this analysis is to identify how race, tumor characteristics, treatment characteristics and social factors impact EC survival.

This analysis uses data from the SEER\*Stat Database: Incidence - SEER Research Plus Data, 17 Registries, Nov 2023 Sub (2000-2021). Non-Hispanic Black and Non-Hispanic White women were included if they were diagnosed between 2000 and 2021 with invasive EC. Patients younger than 18 years, male patients, and patients with a previous history of cancer or any secondary tumor diagnoses were excluded from the analysis.

### Results

#### Table 1:

Patient demographic and clinicopathological characteristics among non-Hispanic white (NHW) and non-Hispanic black (NHB) women diagnosed with endometrial cancer in 2000–2021.

```{r table 1, echo=FALSE}
readRDS(here::here("Output/table_1.rds"))
```

#### Table 2:

Age-adjusted hazard ratios (HRs) and 95% confidence intervals (CIs) for Endometrial Cancer (EC)-specific death according to race among women diagnosed with EC in 2000–2021.

```{r cox model, echo=FALSE}
readRDS(here::here("Output/table_2.rds"))
```

#### Table 3:

Age-adjusted and multivariable-adjusted hazard ratios (HRs) and 95% confidence intervals (CIs) for Endometrial Cancer (EC)-specific death according to race and tumor characteristics among women diagnosed with EC in 2000-2021.

```{r echo=FALSE}
readRDS(here::here("Output/table_3.rds"))
```

#### Figure 1:

Number of women diagnosed with invasive Endometrial Cancer per year in the SEER\*STAT database between 2000-2021.

```{r echo=FALSE}
readRDS(here::here("Output/figure_1.rds"))
```

### Conclusions

Rates of endometrial cancer in the United States has increased substantially since 2000 (Figure 1). While there was a noticeable decrease of cases in 2020, it can be assumed that the COVID-19 pandemic prevent may from accessing healthcare.

Similar to trends in other cancers, Non-Hispanic black women are more likely to die from endometrial cancer than Non-Hispanic white women (Table 2). Non-Hispanic black women are also diagnosed at more advanced stages and with higher grades (Table 1). Rare histologic subtypes, such as Mixed Cell Adenocarcinoma or Clear Cell are more common among Non-Hispanic black women (Table 1). The rare histologic subtypes have worse outcomes than the common endometriod subtypes (Table 3).

